### 2018-06-04 energy_clustering/old_stuff/icml18/rebuttal/rebuttal.txt
energy pami

* R1

We highly appreciate the referee's careful reading of the manuscript 
and comments.

We remark that this paper is the first to relate energy statistics, 
clustering and kernel methods, therefore we believe there is a theoretical 
significance.

The goal of the experiments was never to compete with standard clustering
methods, where a careful tuning of parameters and choice of kernels is
important. Rather, the experiments has the goal of comparing Hartigan's 
method versus Lloyd's method when applied to our QCQP problem, and also
to illustrate the flexibility of energy clustering.
We don't see much relevance in the fact that a scikit-learn implementation of 
spectral clustering, with a careful kernel choice, has a high performance 
in a particular example (we could for instance, introduce a local version
of energy clustering through more sophisticated methods, and we are confident
that this would be better than standard spectral clustering implementations).

* R2

We greatly appreciate the referee's analysis of our paper and comments.

We agree that we should emphasize better the two mentioned contributions.
However, the previous related works cited in the paper, only connects energy
statistics to clustering in a particular case (Euclidean) and there are no
connections to kernels at all, therefore, we believe our paper closes
an important gap and provides a more consistent derivation of clustering 
from energy statistics, in more general settings, and at the same time 
relates this approach to kernel methods.

To the best of our knowledge, Hartigan's method was never applied to
a general kernel method before. We believe this is the advantage of our
approach, where after one chooses a semi-metric, or equivalently, a positive 
definite kernel, one can just run the algorithm proposed in the paper.

We agree that the real data experiment needs a comparison and more detailed
discussion.

* R3

We highly appreciate the referee's careful reading of our manuscript 
and comments.

We kindly disagree with the statement "the contribution of the paper is quite
marginal given the previous work using energy statistics theory for clustering
purpose (e.g, Li (2015) for Euclidean case)". Our paper is the first to 
relate energy statistics, clustering, and kernel methods. Going from Euclidean 
to a general semi-metric with connections with RKHS is not straightforward.
For instance, in going from k-means to kernel k-means there is quite a big
gap and much research involved. The same is true for energy statistics.
It was proposed in the 80's, but only in 2013 connections to kernels were
established, and this was not previously explored in clustering.

We agree that it is desirable to establish the results of Theorems 4 and 5
under general kernels. We believe this is an interesting topic for future
work since it is not straightforward.

The point of Figure 1 is to illustrate that energy clustering attains high 
accuracy, under the same choice of parameters, in different settings. 
Thus, we still believe the comparison in Figure 1 is fair, regardless
if GMM is not the right model to the data.