### 2018-06-04 energy_clustering/old_stuff/code/synapse_data2/Extras/gsheet.Rmd

---
title: ""
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    highlight: pygments
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 2
---
```{r knitOPTS, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, dev = "png")
```

```{r render, eval=FALSE, echo=FALSE}
require(rmarkdown)
#rmarkdown::render("collman14v2_meda_plots.Rmd")
#system("open collman14v2_meda_plots.html")
```

```{r setup,include=FALSE,results='asis',message=FALSE,warning=FALSE, echo = FALSE}
# Library calls here.
require(rmarkdown)
require(knitr)
require(rhdf5)
require(foreach)
require(googlesheets)
suppressMessages(require(meda))
```

## Make plots

```{r, eval = FALSE}
gs_auth(new_user = TRUE)
gs_ls()

tmp <- gs_title("MNSite3Synaptograms.csv")
#dat <- as.data.frame(gs_read_csv(ss = tmp, ws = 1, skip = 1))
dat <- as.data.table(gs_read_csv(ss = tmp, ws = 1, skip = 1))

dat$cx <- 
  (dat$maxX - dat$minX)/2 + dat$minX

dat$cy <- 
  (dat$maxY - dat$minY)/2 + dat$minY

dat$cz <- 
  (dat$maxZ - dat$minZ)/2 + dat$minZ

dat$atx <- round((dat$cx * 3) / 96)
dat$aty <- round((dat$cy * 3) / 96)
dat$atz <- round(dat$cz)

dat$nmx <- (dat$cx / 3)
dat$nmy <- (dat$cy / 3)
dat$nmz <- dat$cz/50

dx <- 438
dy <- 460
dz <- 49

ind <- 
  (dat$atx - 5) >= 0  &
  (dat$aty - 5) >= 0  &
  (dat$atz - 5) >= 0  &
  (dat$atx + 6) <= dx &
  (dat$aty + 6) <= dy &
  (dat$atz + 6) <= dz

dat$buff <- as.numeric(ind)

dat$gaba <- dat$GABA
newDat <- dat[buff == 1, .(atx,aty,atz, id, gaba, shaft, TdTom, NotSynapse)]

write.csv(newDat, file = "GoogleDocData.csv", row.names = FALSE)
write.csv(newDat[, .(atx,aty,atx,id)], file = "GoogleDocData_loc.csv", row.names = FALSE)

inIDS <- as.numeric(read.csv("rorb_avg_at_orderLocations.csv")[[1]])
IDS123 <- as.numeric(read.csv("rorb_avg_at_IDS123.csv")[[1]])

#L <- foreach(i = inIDS, .combine = 'rbind') %do% {
L <- foreach(i = IDS123, .combine = 'rbind') %do% {
  dat[id == i,]
}

glut <- L$glut
glutIDS <- data.table(id = inIDS, glut = as.numeric(glut))
write.csv(glutIDS, file = "rorb_glut.csv", row.names = FALSE)

gaba <- L$gaba
gabaIDS <- data.table(id = inIDS, gaba = as.numeric(gaba))
write.csv(gabaIDS, file = "rorb_gaba.csv", row.names = FALSE)

shaft <- L$shaft
shaftIDS <- data.table(id = inIDS, shaft = as.numeric(shaft))
write.csv(shaftIDS, file = "rorb_shaft.csv", row.names = FALSE)

TdTom <- L$TdTom
TdIDS <- data.table(id = inIDS, TdTom = as.numeric(TdTom))
write.csv(TdIDS, file = "rorb_TdTom.csv", row.names = FALSE)

write.csv(L, file = "rorb_id123subset.csv", row.names = FALSE)
```